\Large Generative Dynamics of Supreme Court Citations

### 2020-04-14 Tex/paper.tex
Started Updating Results-section

\subsection{c-TERGM Results}
The results from the c-TERGM are presented in Table \ref{tab:tergmcoef} and Figures \ref{fig:coeftrends_endo} and \ref{fig:coeftrends_exo}. Table \ref{tab:tergmcoef} gives the coefficient estimates and 95\% bootstrap confidence intervals. In Figures, \ref{fig:coeftrends_endo} and \ref{fig:coeftrends_exo}, we depict coefficient estimates and confidence intervals for all of the effects for which we estimated trends (excluding the intercept---edge count---which is typically not interpreted substantively). The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one unit increase in the respective variable. We first discuss the dependence effect estimates (panels (a)--(d) in Figure \ref{fig:coeftrends_endo}). The reciprocity effect is substantial, positive, and statistically significant for most of the period under study. Conditional upon cases $i$ and $j$ being considered within the same term, the log of the odds that opinion $i$ cites opinion $j$ increases by approximately $2$ when opinion $j$ cites opinion $i$. The naive probability that any one case cites another case decided at the same time or previously in the Supreme Court Citation network is approximately $0.0019$. This probability increases eight-fold, to approximately $0.015$ if the citation would be reciprocal. The transitivity (i.e., triangle closure) effect is also substantial, positive, and statistically significant over the entire period studied. The probability that case $i$ cites case $j$ increases approximately five-fold if the citation would close a triangle involving cases $i$ and $j$. The activity and popularity effects are both estimated to be positive, but trend downward and are more modest in scale (and in the case of activity, not statistically significant later in the time series). Interpreting the popularity effect---the log odds that $i$ cites $j$ increases by 0.018--0.035 with every additional citation to case $j$. Based on the in-degree distribution, it appears that the popularity of cases can be classified by multiples of 10. If case $j$ increases from ten to twenty citations from other cases, the probability that case $i$ cites case $j$ increases from approximately 0.0030 to 0.0040. Overall, these results represent evidence of complex dependence in Supreme Court citation formation.

The effects of the exogenous covariates included in the c-TERGM (panels (e)--(k) in Figure \ref{fig:coeftrends_exo}) exhibit some surprising dynamics. The effects of absolute difference in Martin-Quinn score, and majority size of the recipient case all flip signs throughout the time period under study. This may be an artifact of the simple linear trends we have estimated, but is interesting nonetheless. The attitudinal model holds that ideological preferences drive decision-making on the US Supreme Court. Ideological selection of precedents would be evidenced by a negative effect of the absolute difference in MQ scores in the c-TERGM. We find this effect is negative and statistically significant only after 1975. The absolute difference in MQ scores variable is on the scale of approximately 0---3. At its most ideological point (i.e. 2001), the probability that case $i$ cites case $j$ decreases from approximately 0.024 to 0.018---by one third---if the ideological distance between the majority coalitions in the two cases increases by 3. The effects of the majority size and ideological breadth of case $j$ are similar in both scale and trend. It is not until later in the time series that they both exhibit their expected positive effects. The effect of same issue is positive and statistically significant, and substantial for the time period under study. The probability that $i$ cites $j$ increases by approximately three-fold over the naive probability of citation if $i$ and $j$ have the same legal issue area.

@@

{\em {\bf Note:} The results section of this paper is incomplete and in flux. We recently discovered evidence of non-convergence in the C-TERGM results currently presented in the paper, and have changed the models and estimation settings to address this problem. Most of the substantive conclusions look like they will hold using the updated models, but we are not sure that we will be able to include the updated models in the APSA presentation.}

@@

\begin{table}[htp]
\footnotesize
\centering
\begin{tabular}{|
>{\columncolor[HTML]{EFEFEF}}l |l|l|l|l|} 
\hline
                                                   & \cellcolor[HTML]{EFEFEF}Estimate & \cellcolor[HTML]{EFEFEF}Lower Bound & \cellcolor[HTML]{EFEFEF}Upper Bound & \cellcolor[HTML]{EFEFEF}Significance \\
                                                    \hline
Edges                                              & -6.249                           & -6.412                              & -6.110                              & *                                    \\ \hline
Instar(2)                                          & 0.035                            & 0.032                               & 0.038                               & *                                    \\ \hline
Outstar(2)                                         & 0.021                            & 0.019                               & 0.028                               & *                                    \\ \hline
Mutual                                             & 2.071                            & -4.157                               & 4.072                                &                                     \\ \hline
Triangle                                           & 1.507                            & 1.431                               & 1.559                               & *                                    \\ \hline
Martin Quinn Score                                 & 0.079                            & 0.029                               & 0.130                               & *                                    \\ \hline
Same Issue Area Homophily                          & 1.447                            & 1.389                               & 1.515                               & *                                    \\ \hline
Year Difference                                    & -0.068                           & -0.079                              & -0.058                              & *                                    \\ \hline
$(\text{Year Differnce})^2$                        & 0.0019                           & 0.0016                              & 0.0022                              & *                                    \\ \hline
Receiver Abs Diff MQ Score in Majority             & 0.047                           & 0.029                              & 0.063                              & *                                     \\ \hline
Receiver Number Justices in Majority               & -0.075                           & -0.095                              & -0.055                              & *                                    \\ \hline
Receiver  Sender Year                              & 0.0039                           & 0.0036                              & 0.0042                              & *                                    \\ \hline
Overruled Cases   & -1.043                           & -1.624                              & -0.564                              & *                                    \\ \hline
Sender Same Issue Area 2                           & 0.142                            & 0.100                               & 0.185                               & *                                    \\ \hline
Sender Same Issue Area 3                           & -0.337                           & -0.403                              & -0.270                              & *                                    \\ \hline
Sender Same Issue Area 4                           & 0.498                            & 0.438                               & 0.561                               & *                                    \\ \hline
Sender Same Issue Area 5                           & 0.182                            & 0.031                               & 0.305                               & *                                    \\ \hline
Sender Same Issue Area 6                           & 0.536                            & 0.422                               & 0.648                               & *                                    \\ \hline
Sender Same Issue Area 7                           & 0.370                            & 0.316                               & 0.426                               & *                                    \\ \hline
Sender Same Issue Area 8                           & 0.107                            & 0.059                               & 0.156                               & *                                    \\ \hline
Sender Same Issue Area 9                           & 0.278                            & 0.238                               & 0.329                               & *                                    \\ \hline
Sender Same Issue Area 10                          & 0.371                            & 0.317                               & 0.433                               & *                                    \\ \hline
Sender Same Issue Area 11                          & -0.030                           & -0.151                              & 0.164                               &                                      \\ \hline
Sender Same Issue Area 12                          & 0.064                           & -0.014                              & 0.130                               &                                      \\ \hline
Sender Same Issue Area 13                          & 0.608                            & 0.423                               & 0.792                               & *                                    \\ \hline
Sender Same Issue Area 14                          & 0.232                            & 0.075                              & 0.393                               & *                                      \\ \hline
Receiver Same Issue Area 2                         & 0.252                            & 0.199                               & 0.310                               & *                                    \\ \hline
Receiver Same Issue Area 3                         & -0.092                           & -0.194                              & 0.014                              &                                     \\ \hline
Receiver Same Issue Area 4                         & 0.481                            & 0.397                               & 0.570                               & *                                    \\ \hline
Receiver Same Issue Area 5                         & 0.374                            & 0.224                               & 0.519                               & *                                    \\ \hline
Receiver Same Issue Area 6                         & 0.521                            & 0.346                               & 0.652                               & *                                    \\ \hline
Receiver Same Issue Area 7                         & 0.412                            & 0.339                               & 0.487                               & *                                    \\ \hline
Receiver Same Issue Area 8                         & 0.174                            & 0.134                               & 0.232                               & *                                    \\ \hline
Receiver Same Issue Area 9                         & 0.311                            & 0.240                               & 0.361                               & *                                    \\ \hline
Receiver Same Issue Area 10                        & 0.506                            & 0.429                               & 0.577                               & *                                    \\ \hline
Receiver Same Issue Area 11                        & 0.299                            & 0.088                               & 0.432                               & *                                    \\ \hline
Receiver Same Issue Area 12                        & 0.350                            & 0.276                               & 0.445                               & *                                    \\ \hline
Receiver Same Issue Area 13                        & 0.972                            & 0.780                               & 1.135                               & *                                    \\ \hline
Receiver Same Issue Area 14                        & 0.517                            & 0.200                               & 0.740                               & *                                    \\ \hline
Instar(2) $\times$ Sender Year                     & -0.00026                         & -0.00031                            & -0.00021                            & *                                    \\ \hline
Outstar(2) $\times$ Sender Year                    & -0.00031                         & -0.00045                            & -0.00027                            & *                                    \\ \hline
Mutual $\times$ Sender Year                        & 0.110                           & 0.010                              & 0.524                               & *                                      \\ \hline
Triangle $\times$ Sender Year                      & -0.00022                          & -0.0016                             & 0.0013                              &                                      \\ \hline
Martin Quinn Score $\times$ Sender Year            & -0.0019                          & -0.0030                             & -0.0009                             & *                                    \\ \hline
Same Issue Area $\times$ Sender Year               & -0.0068                          & -0.0081                             & -0.0055                             & *                                    \\ \hline
Year Difference $\times$ Sender Year               & 0.00046                           & 0.00028                              & 0.00065                              & *                                    \\ \hline
$\text{Year Difference}^2 \times$ Sender Year      & -0.000024                        & -0.000028                           & -0.000021                           & *                                    \\ \hline
MQ Score in Majority $\times$ Sender Year & -0.00053                           & -0.00093                              & -0.0018                              & *                                    \\ \hline
Justices in Majority $\times$ Sender Year   & 0.0017                           & 0.0013                              & 0.0021                              & *                                    \\ \hline
Overruled Cases $\times$ Sender Year   & 0.012                           & 0.002                              & 0.023                              & *                                    \\ \hline
\end{tabular}
\caption{Bootstrapped MPLE Results for the time period $1937-2015$. A '*' indicates that the 2.5th and 97.5th quantile of the variable does not include '0' and as a result is statistically significant.}
\label{tab:tergmcoef}
\end{table}

>>>>

\subsection{c-TERGM Results}
The results from the c-TERGM are presented Figures \ref{SCC_results_1} and \ref{SCC_results_2}. In these figures we depict coefficient estimates and 95\% confidence intervals for all of the effects for all the variables we discussed in sections \ref{covariate_terms} and \ref{dependence_terms}. The coefficients in an ERGM family model can be interpreted in the same way as logistic regression coefficients---they give the change in the log odds of a tie from $i$ to $j$ given a one unit increase in the respective variable. For this reason, the number of edges in a network corresponds to the intercept in an ERGM. We first discuss the dependence effect estimates in Figure \ref{SCC_results_1}. A circle indicates that the estimate was significant at an $\alpha=0.05$ level, while and square translates to a p-value between $0.05$ and $0.1$. A triangle signals that the estimate for a given year was not significant at an $\alpha=0.1$ level.

%The reciprocity effect is substantial, positive, and statistically significant for most of the period under study. Conditional upon cases $i$ and $j$ being considered within the same term, the log of the odds that opinion $i$ cites opinion $j$ increases by approximately $2$ when opinion $j$ cites opinion $i$. The naive probability that any one case cites another case decided at the same time or previously in the Supreme Court Citation network is approximately $0.0019$. This probability increases eight-fold, to approximately $0.015$ if the citation would be reciprocal. The transitivity (i.e., triangle closure) effect is also substantial, positive, and statistically significant over the entire period studied. The probability that case $i$ cites case $j$ increases approximately five-fold if the citation would close a triangle involving cases $i$ and $j$. The activity and popularity effects are both estimated to be positive, but trend downward and are more modest in scale (and in the case of activity, not statistically significant later in the time series). Interpreting the popularity effect---the log odds that $i$ cites $j$ increases by 0.018--0.035 with every additional citation to case $j$. Based on the in-degree distribution, it appears that the popularity of cases can be classified by multiples of 10. If case $j$ increases from ten to twenty citations from other cases, the probability that case $i$ cites case $j$ increases from approximately 0.0030 to 0.0040. Overall, these results represent evidence of complex dependence in Supreme Court citation formation.

%The effects of the exogenous covariates included in the c-TERGM (panels (e)--(k) in Figure \ref{fig:coeftrends_exo}) exhibit some surprising dynamics. The effects of absolute difference in Martin-Quinn score, and majority size of the recipient case all flip signs throughout the time period under study. This may be an artifact of the simple linear trends we have estimated, but is interesting nonetheless. The attitudinal model holds that ideological preferences drive decision-making on the US Supreme Court. Ideological selection of precedents would be evidenced by a negative effect of the absolute difference in MQ scores in the c-TERGM. We find this effect is negative and statistically significant only after 1975. The absolute difference in MQ scores variable is on the scale of approximately 0---3. At its most ideological point (i.e. 2001), the probability that case $i$ cites case $j$ decreases from approximately 0.024 to 0.018---by one third---if the ideological distance between the majority coalitions in the two cases increases by 3. The effects of the majority size and ideological breadth of case $j$ are similar in both scale and trend. It is not until later in the time series that they both exhibit their expected positive effects. The effect of same issue is positive and statistically significant, and substantial for the time period under study. The probability that $i$ cites $j$ increases by approximately three-fold over the naive probability of citation if $i$ and $j$ have the same legal issue area.

@@

%{\em {\bf Note:} The results section of this paper is incomplete and in flux. We recently discovered evidence of non-convergence in the C-TERGM results currently presented in the paper, and have changed the models and estimation settings to address this problem. Most of the substantive conclusions look like they will hold using the updated models, but we are not sure that we will be able to include the updated models in the APSA presentation.}

@@
